---
title: "Pain Research - Dr. Jamison Data"
author: "Jiaqi Bi"
date: "2023-09-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Two different files
#setwd("~/Desktop/Project; Daily diaries; 30-days/Pain-Research-1-Dr-Marc-O-Martel/Datasets; Project; Martel")
#setwd("~/Desktop/Project; Daily diaries; 30-days/Pain-Research-1-Dr-Marc-O-Martel/Datasets; Project; Jamison")
```

```{r, include = FALSE, eval = TRUE, message = FALSE}
## Load packages
library(tidyverse)
library(ggplot2)
library(tidyr)
library(haven) ## This library provides functions to read sav file into R
library(lme4)
library(lmerTest)
library(performance)
```

```{r, include = FALSE}
## Baseline data
#df1 <- read_sav("Dataset; Jamison project; Baseline.sav")
#df1 <- read_sav("Dataset; Jamison project; Baseline_Excluded.sav")
#df1 <- read_sav("E:/UWO/DR Marc O Martel data/Pain-Research-1-Dr-Marc-O-Martel/Datasets; Project; Jamison/Dataset; Jamison project; Baseline.sav")
## Daily data
#df2 <- read_sav("Dataset; Jamison project; Daily.sav")
#df2 <- read_sav("E:/UWO/DR Marc O Martel data/Pain-Research-1-Dr-Marc-O-Martel/Datasets; Project; Jamison/Dataset; Jamison project; Daily.sav")



## Examine the duplicated observation
#df1$ID[which(duplicated(df1$ID))]
#df2$StudyID[which(duplicated(df2$StudyID))]

#df1 <- df1 |>
#  filter(!duplicated(ID))

#df1 <- df1 |> filter(Exclud_MissingBoth2 == 0)
#df2 <- df2 |>
#  filter(!duplicated(StudyID)) |>
#  rename(ID = StudyID)

## Time variable to numerical day (consecutive)
#date_cols <- grep("^Date", names(df2), value = TRUE)

#df2 <- df2 %>%
#  mutate(across(all_of(date_cols), as.Date, format = "%Y-%m-%d"))

## Correct abberrant years
#df2 <- df2 %>%
#  rowwise() %>%
#  mutate(across(all_of(date_cols), ~ {
#    if (. != Date1 && !is.na(.)) {
#      day_diff <- as.numeric(. - Date1)
#      if (day_diff < 0) {
#        update(., year = year(Date1))
#      } else {
#        .
#      }
#    } else {
 #     .
#    }
#  })) %>%
#  ungroup()

####### Below investigates the rest of aberrant objects #######
##### Above does not solve those entering a new year ##########
#df2_investigate <- df2 |>
#  pivot_longer(
#  cols = starts_with("Date"), 
#  names_to = "Date_Number", 
#  values_to = "Date_Value"
#) |>
#  select(ID, Date_Value)

#print(n = 36, df2_investigate[df2_investigate$ID == 673,])
###############################################################

## Below solved all objects date problem
#df2$Date8[df2$ID == 610] <- "2023-01-01"
#df2$Date8[df2$ID == 703] <- "2022-01-09"
#df2$Date8[df2$ID == 741] <- "2021-01-10"
#df2$Date8[df2$ID == 680] <- "2023-01-09"
#df2$Date28[df2$ID == 673] <- "2023-02-02"

## Convert the date to number
#df2 <- df2 |>
#  mutate(across(all_of(date_cols[-1]), ~ as.numeric(. - df2$Date1 + 1)))

#df2$Date1 <- 1
#all_cols <- names(df2)

#for (i in 2:length(date_cols)) {
#  if (any(df2[[date_cols[i]]] == 1, na.rm = TRUE)) {
#    same_day_cols <- grep(paste0("_", i), all_cols, value = TRUE)
#    df2[df2[[date_cols[i]]] == 1 & !is.na(df2_test[[date_cols[i]]]), same_day_cols] <- NA
#  }
#}

## Wide to long
#df2_long <- df2 |>
#  pivot_longer(cols = -ID,
#               names_to = c(".value", "day"),
#               names_pattern = "([A-Za-z]+)(\\d+)") |>
#  select(-day) |>
#  rename(Day = Date) |>
#  filter(Day <= 30 & !is.na(Day)) |>
#  group_by(ID) |>
#  distinct(Day, .keep_all = TRUE) |>
#  complete(Day = seq(1,30))

## replace all -1 to NA
#df2_long <- df2_long |>
#  mutate_all(~if_else(. < 0, NA_real_, .))

#df2_long <- df2_long |>
#  rename(GlobalImprovement = Changed)

## Calculate the lagged variable
#df2_long <- df2_long |>
#  group_by(ID) |>
#  arrange(ID, Day) |>
#  mutate(AvePain_Change = AvePain - lag(AvePain), 
#         ActivityInt_Change = ActivityInt - lag(ActivityInt), 
#         Mood_Change = Mood - lag(Mood)) 

#df2_greaterthan7 <- df2_long |>
#  group_by(ID) %>%
#  summarise(NonMissingGI = sum(!is.na(GlobalImprovement))) %>%
#  filter(NonMissingGI >= 7) %>%
#  inner_join(df2_long, by = "ID")

## Add level 2 age, gender, and PCS to the longitudinal data
#df1_temporary <- df1 |>
#  select(c(B_Demog_Age, B_Demog_Gender, B_Psych_PCS, ID))
#df1_temporary <- df1

#df_new <- merge(df2_greaterthan7, df1_temporary, by = "ID")
#df_total <- merge(df2_greaterthan7, df1, by = "ID")

## Global improvement 10 to 1, 9 to 2, ...
#df_new <- df_new |>
#  mutate(GlobalImprovement = 11-GlobalImprovement)

#df_new <- df_new |> filter(ID %in% df1$ID)
#write_sav(df_new, "Data; Jamison data; Long format (Stacked).sav")
df_new <- read_sav("File To; Jiaqi; 2024.09.2.sav")
#df_new <- df_new |>
#  group_by(ID) |>
#  mutate(across(c(AvePain, Sleep, Mood, ActivityInt), ~. - mean(., na.rm = TRUE),
#                .names = "{.col}_c"))

group_means <- aggregate(cbind(AvePain, Sleep, Mood, ActivityInt) ~ ID, data = df_new, FUN = function(x) mean(x, na.rm = TRUE))

df_new <- merge(df_new, group_means, by = "ID", suffixes = c("", "_mean"))

df_new$AvePain_c <- df_new$AvePain - df_new$AvePain_mean
df_new$Sleep_c <- df_new$Sleep - df_new$Sleep_mean
df_new$Mood_c <- df_new$Mood - df_new$Mood_mean
df_new$ActivityInt_c <- df_new$ActivityInt - df_new$ActivityInt_mean

df_new <- df_new[, !names(df_new) %in% c("AvePain_mean", "Sleep_mean", "Mood_mean", "ActivityInt_mean")]


#df_new <- df_new %>%
#  mutate(
#    across(
#      c(AvePain, Sleep, Mood, ActivityInt),
#      ~ . - mean(., na.rm = TRUE),
#      .names = "{.col}_c"
#    )
#  )


df_new <- df_new |>
  group_by(ID) |>
  mutate(LastDay_PainAve_c = lag(AvePain_c), 
         NextDay_PainAve_c = lead(AvePain_c), 
         Today_PainAve_c = AvePain_c,
         LastDay_Sleep_c = Sleep_c,
         Today_Sleep_c = lead(Sleep_c),
         LastDay_Mood_c = lag(Mood_c),
         Today_Mood_c = Mood_c,
         Today_PainAve = AvePain,
         Today_Sleep = lead(Sleep),
         LastDay_Sleep = Sleep,
         NextDay_Sleep = lag(Sleep),
         LastDay_PainAve = lag(AvePain)
         )

```

## Objective 1.1 Day-to-day (Concurrent) associations

```{r}
#Analyses; Day-to-day; Univariate multilevel linear regressions 
#- Outcome: Lev1 daily pain intensity 
#- Examine Lev1 association; between daily mood and pain (Same-day Lev1 units) 
#- Examine Lev1 association; between daily sleep and pain (Same-day Lev1 units) 
#- All these multlev must be done with Lev1 centered data 
model_Pain1 <- lmer(Today_PainAve ~ Today_Mood_c  + (1|ID), data = df_new)
summary(model_Pain1)
confint(model_Pain1)
model_Pain2 <- lmer(Today_PainAve ~ Today_Sleep_c + (1|ID), data = df_new)
summary(model_Pain2)
confint(model_Pain2)

# Analyses; Multivariable models   
#- Outcome: Lev1 daily pain intensity 
#- Ivs entered simulatenously: daily (Lev1) mood, sleep 
#- All these multlev must be done with Lev1 centered data 
#- Same-day Lev1 units 
model_Pain3 <- lmer(Today_PainAve ~ Today_Sleep_c + Today_Mood_c + (1|ID), data = df_new)
summary(model_Pain3)
confint(model_Pain3)

# Analyses; Day-to-day; Univariate multilevel linear regressions 
#-Outcome: Lev1 Sleep 
#-Examine Lev1 association; between daily mood and sleep (Same-day Lev1 units) 
#-Examine Lev1 association; between daily pain and sleep (Same-day Lev1 units) 
#-All these multlev must be done with Lev1 centered data 
model_Pain4 <- lmer(Today_Sleep ~ Today_Mood_c + (1|ID), data = df_new)
summary(model_Pain4)
confint(model_Pain4)
model_Pain5 <- lmer(Today_Sleep ~ Today_PainAve_c + (1|ID), data = df_new)
summary(model_Pain5)
confint(model_Pain5)

# Analyses; Multivariable models   
#- Outcome: Lev1 sleep 
#- Ivs entered simulatenously: daily (Lev1) mood, pain 
#- All these multlev must be done with Lev1 centered data 
#- Same-day Lev1 units 
model_Pain6 <- lmer(Today_Sleep ~ Today_PainAve_c + Today_Mood_c + (1|ID), data = df_new)
summary(model_Pain6)
confint(model_Pain6)
```

## Objective 1.2: Time-lag effects 

```{r}
# Yesterday sleep -> Toady Pain
# No lag needed here - LastDay_Sleep_c is equivalent to Sleep_c
model_SleepPain <- lmer(Today_PainAve ~ LastDay_Sleep_c + (1|ID), data = df_new)
icc(model_SleepPain)
summary(model_SleepPain)
confint(model_SleepPain)
# Yesterday pain -> Today sleep
# Today_Sleep is computed using Next Day's Sleep reportings, 
# LastDay_PainAve_c is computed using Last Day's PainAve_c reportings
model_PainSleep <- lmer(Today_Sleep ~ LastDay_PainAve_c + (1|ID), data = df_new)
summary(model_PainSleep)
confint(model_PainSleep)
## Both are within cluster/within individual centered
```

## Objective 2: Anayses: Perceived Improvement

```{r}
# Analyses; Univariate multilevel linear regressions 
# Outcome: Lev1 perceived improvement 
# Examine Lev1 association; between daily pain and perceived improvement (Same-day Lev1 units) 
model_painimprove <- lmer(GlobalImprovement ~ AvePain_c + (1|ID), data = df_new)
summary(model_painimprove)
confint(model_painimprove)
# Examine Lev1 association; between daily mood and perceived improvement (Same-day Lev1 units) 
model_moodimprove <- lmer(GlobalImprovement ~ Mood_c + (1|ID), data = df_new)
summary(model_moodimprove)
confint(model_moodimprove)
# Examine Lev1 association; between daily sleep and perceived improvement (Same-day Lev1 unit) 
model_sleepimprove <- lmer(GlobalImprovement ~ Sleep_c + (1|ID), data = df_new)
summary(model_sleepimprove)
confint(model_sleepimprove)
# Examine Lev1 association; between daily ActivityInt and perceived improvement (Same-day Lev1 unit)
model_actimprove <- lmer(GlobalImprovement ~ ActivityInt_c + (1|ID), data = df_new)
summary(model_actimprove)
confint(model_actimprove)
# All these multlev must be done with Lev1 centered data 

```

```{r}
# Analysis; Multvariable/multilevel linear regression 
# Outcome: Perceived improvement 
# Ivs entered simulatenously: daily (Lev1) pain, mood, sleep, ActivityInt  
model_compimprove <- lmer(GlobalImprovement ~ Sleep_c + AvePain_c + Mood_c + ActivityInt_c + (1|ID), data = df_new)
summary(model_compimprove)
confint(model_compimprove)
# All these multlev must be done with Lev1 centered data 
# Perhaps get some colinearity indicator to know to what extent colinearity is an issue 
collinear_test <- check_collinearity(model_compimprove)
print(collinear_test)
## Some comments on how to read this result: The VIF is around 1 => Low Multicollinearity
## The VIF is between 2 to 5, Moderate Multicollinearity
## The VIF >5 (or 10 sometimes), High Multicollinearity
## Low Tolerance (~ 0), High Multicollinearity
```

We have low multicollinearity in this case!

## Objective 2.2 Analyses: Moderators of perceived impovement

```{r}
# Test if any of the baseline (Lev2) socio-demog variables are linked to perceived improvements; Univariate
# B_Demog_Gender 
model_genderimprove <- lmer(GlobalImprovement ~ B_Demog_Gender + (1|ID), data = df_new)
summary(model_genderimprove)
confint(model_genderimprove)
# B_Demog_Ethnicity 
model_ethnimprove <- lmer(GlobalImprovement ~ B_Demog_Ethnicity + (1|ID), data = df_new)
summary(model_ethnimprove)
confint(model_ethnimprove)
# B_Demog_Age
model_ageimprove <- lmer(GlobalImprovement ~ B_Demog_Age + (1|ID), data = df_new)
summary(model_ageimprove) # Significant! 
confint(model_ageimprove)
```

Age tends to be associated with the improvement, older patients have better improvements. 

```{r}
# Test if any of the baseline (Lev2) clinical variables are linked to perceived improvements; Univariate 
# B_Clin_PainDur"
model_paindurimprove <- lmer(GlobalImprovement ~ B_Clin_PainDur + (1|ID), data = df_new)
summary(model_paindurimprove)
confint(model_paindurimprove)
# B_Clin_BMI"
model_BMIimprove <- lmer(GlobalImprovement ~ B_Clin_BMI + (1|ID), data = df_new)
summary(model_BMIimprove)
confint(model_BMIimprove)
# All the medications below; 
# separately/independently; association with perceived improivement; Univariate 
# B_Med_Tramadol 
model_Tramadolimprove <- lmer(GlobalImprovement ~ B_Med_Tramadol + (1|ID), data = df_new)
summary(model_Tramadolimprove) # Significant!
confint(model_Tramadolimprove)
# B_Med_Suboxone 
model_Suboxoneimprove <- lmer(GlobalImprovement ~ B_Med_Suboxone + (1|ID), data = df_new)
summary(model_Suboxoneimprove)
confint(model_Suboxoneimprove)
# B_Med_Marijuana 
model_Marijuanaimprove <- lmer(GlobalImprovement ~ B_Med_Marijuana + (1|ID), data = df_new)
summary(model_Marijuanaimprove)
confint(model_Marijuanaimprove)
# B_Med_NSAIDS 
model_NSAIDSimprove <- lmer(GlobalImprovement ~ B_Med_NSAIDS + (1|ID), data = df_new)
summary(model_NSAIDSimprove)
confint(model_NSAIDSimprove)
# B_Med_Anticonvulsant 
model_Anticonimprove <- lmer(GlobalImprovement ~ B_Med_Anticonvulsant + (1|ID), data = df_new)
summary(model_Anticonimprove) # Significant!
confint(model_Anticonimprove)
# B_Med_MuscleRelaxer 
model_MuscleRelimprove <- lmer(GlobalImprovement ~ B_Med_MuscleRelaxer + (1|ID), data = df_new)
summary(model_MuscleRelimprove)
confint(model_MuscleRelimprove)
# B_Med_Antidepressants 
model_Antidepimprove <- lmer(GlobalImprovement ~ B_Med_Antidepressants + (1|ID), data = df_new)
summary(model_Antidepimprove)
confint(model_Antidepimprove)
# B_Med_Benzodiazepine 
model_Benzoimprove <- lmer(GlobalImprovement ~ B_Med_Benzodiazepine + (1|ID), data = df_new)
summary(model_Benzoimprove)
confint(model_Benzoimprove)
# B_Med_Stimulants 
model_Stimuimprove <- lmer(GlobalImprovement ~ B_Med_Stimulants + (1|ID), data = df_new)
summary(model_Stimuimprove)
confint(model_Stimuimprove)
# B_Med_OtherMed 
model_OtherMedimprove <- lmer(GlobalImprovement ~ B_Med_OtherMed + (1|ID), data = df_new)
summary(model_OtherMedimprove)
confint(model_OtherMedimprove)
# B_Med_OTC 
model_OTCimprove <- lmer(GlobalImprovement ~ B_Med_OTC + (1|ID), data = df_new)
summary(model_OTCimprove)
confint(model_OTCimprove)
# B_Med_OpioidsYN
model_Opioid_improve <- lmer(GlobalImprovement ~ B_Med_OpioidsYN + (1|ID), data = df_new)
summary(model_Opioid_improve) # Significant! 
confint(model_Opioid_improve)
 
```

B_Med_Tramadol, B_Med_Anticonvulsant, and B_Med_OpioidsYN are associated with the Global Improvement! 

```{r}
# Test if any of the baseline (Lev2) psych variables are linked to perceived improvements; Univariate 
# B_Psych_PCS"
model_PCSimprove <- lmer(GlobalImprovement ~ B_Psych_PCS + (1|ID), data = df_new)
summary(model_PCSimprove) # Significant!
confint(model_PCSimprove)
# B_Psych_HADS"
model_HADSimprove <- lmer(GlobalImprovement ~ B_Psych_HADS + (1|ID), data = df_new)
summary(model_HADSimprove) # Significant! 
confint(model_HADSimprove)

```

Both are significant! 

```{r}
# Then if any of the Lev2 variables above are significantly associated with the outcome (i.e., perceived improvement), then we’ll test the cross-level (Lev2 X Lev1) interaction effects to see if these Lev2 variables moderate the effects of : 
#1; Sleep on daily improvement 
model_ageimprove_Sleep <- lmer(GlobalImprovement ~ B_Demog_Age*Sleep_c + (1|ID), data = df_new)
summary(model_ageimprove_Sleep) 
confint(model_ageimprove_Sleep)
model_Tramadolimprove_Sleep <- lmer(GlobalImprovement ~ B_Med_Tramadol*Sleep_c + (1|ID), data = df_new)
summary(model_Tramadolimprove_Sleep) 
confint(model_Tramadolimprove_Sleep)
model_Anticonimprove_Sleep <- lmer(GlobalImprovement ~ B_Med_Anticonvulsant*Sleep_c + (1|ID), data = df_new)
summary(model_Anticonimprove_Sleep) 
confint(model_Anticonimprove_Sleep)
model_Opioid_improve_Sleep <- lmer(GlobalImprovement ~ B_Med_OpioidsYN*Sleep_c + (1|ID), data = df_new)
summary(model_Opioid_improve_Sleep) 
confint(model_Opioid_improve_Sleep)
model_PCSimprove_Sleep <- lmer(GlobalImprovement ~ B_Psych_PCS*Sleep_c + (1|ID), data = df_new)
summary(model_PCSimprove_Sleep) 
confint(model_PCSimprove_Sleep)
model_HADSimprove_Sleep <- lmer(GlobalImprovement ~ B_Psych_HADS*Sleep_c + (1|ID), data = df_new)
summary(model_HADSimprove_Sleep) 
confint(model_HADSimprove_Sleep)
#2; Mood on daily improvement 
model_ageimprove_Mood <- lmer(GlobalImprovement ~ B_Demog_Age*Mood_c + (1|ID), data = df_new)
summary(model_ageimprove_Mood) 
confint(model_ageimprove_Mood)
model_Tramadolimprove_Mood <- lmer(GlobalImprovement ~ B_Med_Tramadol*Mood_c + (1|ID), data = df_new)
summary(model_Tramadolimprove_Mood) 
confint(model_Tramadolimprove_Mood)
model_Anticonimprove_Mood <- lmer(GlobalImprovement ~ B_Med_Anticonvulsant*Mood_c + (1|ID), data = df_new)
summary(model_Anticonimprove_Mood) 
confint(model_Anticonimprove_Mood)
model_Opioid_improve_Mood <- lmer(GlobalImprovement ~ B_Med_OpioidsYN*Mood_c + (1|ID), data = df_new)
summary(model_Opioid_improve_Mood) 
confint(model_Opioid_improve_Mood)
model_PCSimprove_Mood <- lmer(GlobalImprovement ~ B_Psych_PCS*Mood_c + (1|ID), data = df_new)
summary(model_PCSimprove_Mood) 
confint(model_PCSimprove_Mood)
model_HADSimprove_Mood <- lmer(GlobalImprovement ~ B_Psych_HADS*Mood_c + (1|ID), data = df_new)
summary(model_HADSimprove_Mood) 
confint(model_HADSimprove_Mood)
#3; Pain on daily improvement 
model_ageimprove_AvePain <- lmer(GlobalImprovement ~ B_Demog_Age*AvePain_c + (1|ID), data = df_new)
summary(model_ageimprove_AvePain) 
confint(model_ageimprove_AvePain)
model_Tramadolimprove_AvePain <- lmer(GlobalImprovement ~ B_Med_Tramadol*AvePain_c + (1|ID), data = df_new)
summary(model_Tramadolimprove_AvePain) 
confint(model_Tramadolimprove_AvePain)
model_Anticonimprove_AvePain <- lmer(GlobalImprovement ~ B_Med_Anticonvulsant*AvePain_c + (1|ID), data = df_new)
summary(model_Anticonimprove_AvePain) 
confint(model_Anticonimprove_AvePain)
model_Opioid_improve_AvePain <- lmer(GlobalImprovement ~ B_Med_OpioidsYN*AvePain_c + (1|ID), data = df_new)
summary(model_Opioid_improve_AvePain) 
confint(model_Opioid_improve_AvePain)
model_PCSimprove_AvePain <- lmer(GlobalImprovement ~ B_Psych_PCS*AvePain_c + (1|ID), data = df_new)
summary(model_PCSimprove_AvePain) 
confint(model_PCSimprove_AvePain)
model_HADSimprove_AvePain <- lmer(GlobalImprovement ~ B_Psych_HADS*AvePain_c + (1|ID), data = df_new)
summary(model_HADSimprove_AvePain) 
confint(model_HADSimprove_AvePain)
#4; ActivInterf on daily improvement 
model_ageimprove_ActivityInt <- lmer(GlobalImprovement ~ B_Demog_Age*ActivityInt_c + (1|ID), data = df_new)
summary(model_ageimprove_ActivityInt) 
confint(model_ageimprove_ActivityInt)
model_Tramadolimprove_ActivityInt <- lmer(GlobalImprovement ~ B_Med_Tramadol*ActivityInt_c + (1|ID), data = df_new)
summary(model_Tramadolimprove_ActivityInt) 
confint(model_Tramadolimprove_ActivityInt)
model_Anticonimprove_ActivityInt <- lmer(GlobalImprovement ~ B_Med_Anticonvulsant*ActivityInt_c + (1|ID), data = df_new)
summary(model_Anticonimprove_ActivityInt) 
confint(model_Anticonimprove_ActivityInt)
model_Opioid_improve_ActivityInt <- lmer(GlobalImprovement ~ B_Med_OpioidsYN*ActivityInt_c + (1|ID), data = df_new)
summary(model_Opioid_improve_ActivityInt) 
confint(model_Opioid_improve_ActivityInt)
model_PCSimprove_ActivityInt <- lmer(GlobalImprovement ~ B_Psych_PCS*ActivityInt_c + (1|ID), data = df_new)
summary(model_PCSimprove_ActivityInt) 
confint(model_PCSimprove_ActivityInt)
model_HADSimprove_ActivityInt <- lmer(GlobalImprovement ~ B_Psych_HADS*ActivityInt_c + (1|ID), data = df_new)
summary(model_HADSimprove_ActivityInt) 
confint(model_HADSimprove_ActivityInt)
```

## Graphs

```{r}
# Graphs; Objective 1 
# Obj1: Graph; Descriptive; Violon plots for 30-day aggregates; Sleep; Mood; Pain; ActivityInt; 
# All on the same 0-10 scale; follow the order listed here.  
```

```{r, echo = FALSE}
data_forgraph <- df_new |>
  group_by(ID) |>
  summarise(AggSleep = mean(Sleep, na.rm = TRUE),
            AggMood = mean(Mood, na.rm = TRUE),
            AggPain = mean(AvePain, na.rm = TRUE), 
            AggActivityInt = mean(ActivityInt, na.rm = TRUE))
data_forgraph <- data_forgraph |>
  pivot_longer(cols = c(AggSleep, AggMood, AggPain, AggActivityInt),
               names_to = "Variables",
               values_to = "Score")
data_forgraph |>
  ggplot(aes(x = Variables, y = Score)) +
  geom_violin(aes(fill = Variables), trim = FALSE, alpha = 0.7) +
  geom_jitter(size = 1) +
  theme_minimal() +
  scale_y_continuous(
    limits = c(0, 10),
    breaks = seq(0, 10, by = 1),
    oob = scales::squish  
  ) +
  labs(
    title = "Violin Plot of Aggregated Variables",
    x = "Variables",
    y = "Scores"
  ) +
  theme(
    legend.position = "none",
    text = element_text(size = 12),
    plot.title = element_text(
      size = 12,                      
      hjust = 0.45,                    
      margin = margin(b = 20)         
    ),
    axis.title.x = element_text(
      margin = margin(t = 15)         
    ),
    axis.title.y = element_text(
      margin = margin(r = 15)         
    ),
    axis.text.x = element_text(
      margin = margin(t = 10)         
    ),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5)
  )

```

```{r}
# Obj1: Graph; Multilevel linear reg slopes; Lagged effects; (1) Sleep/IV & Pain 
```

```{r, echo = FALSE}
fixed_effect <- fixef(model_SleepPain)
fixed_intercept <- fixed_effect["(Intercept)"]
fixed_slope <- fixed_effect["LastDay_Sleep_c"] ## Fixed effects

random_effects <- ranef(model_SleepPain)$ID
random_effects_df <- data.frame(
  ID = rownames(random_effects),
  RandomIntercept = random_effects[, "(Intercept)"]
) ## Random effects

random_lines <- random_effects_df %>%
  mutate(
    Intercept = fixed_intercept + RandomIntercept,
    Slope = fixed_slope
  ) ## Random intercept lines prep

sleep_range <- seq(
  from = min(df_new$LastDay_Sleep_c, na.rm = TRUE),
  to = max(df_new$LastDay_Sleep_c, na.rm = TRUE),
  length.out = 100
)  ## Define the range of predictors

prediction_df <- random_lines %>%
  crossing(LastDay_Sleep_c = sleep_range) %>%
  mutate(Today_PainAve = Intercept + Slope * LastDay_Sleep_c)

fixed_line_df <- data.frame(
  LastDay_Sleep_c = sleep_range,
  Today_PainAve = fixed_intercept + fixed_slope * sleep_range
)

ggplot() +
  geom_line(
    data = prediction_df,
    aes(x = LastDay_Sleep_c, y = Today_PainAve, group = ID),
    color = "grey",
    linetype = "dashed",
    alpha = 0.5
  ) +
  geom_line(
    data = fixed_line_df,
    aes(x = LastDay_Sleep_c, y = Today_PainAve),
    color = "black",
    size = 1
  ) +
  scale_y_continuous(
    limits = c(0, 10),
    breaks = seq(0, 10, by = 1),
    oob = scales::squish  
  ) +
  scale_x_continuous(
    limits = c(min(floor(df_new$LastDay_Sleep_c), na.rm = TRUE), 
               max(ceiling(df_new$LastDay_Sleep_c), na.rm = TRUE)),
    breaks = seq(min(floor(df_new$LastDay_Sleep_c), na.rm = TRUE), 
                 max(ceiling(df_new$LastDay_Sleep_c), na.rm = TRUE), by = 1),
    oob = scales::squish  
  ) +
  theme_minimal() +
  labs(
    title = "Random Intercept Plot: Sleep vs. Pain",
    x = "Last Day Sleep (Centered)",
    y = "Today Average Pain"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(
      size = 12,                      
      hjust = 0.45,                    
      margin = margin(b = 20)         
    ),
    axis.title.x = element_text(
      margin = margin(t = 15)         
    ),
    axis.title.y = element_text(
      margin = margin(r = 15)         
    ),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5)
  )
```

```{r}
# Obj1: Graph; Multilevel linear reg slopes; Lagged effects; (2) Pain/IV & Sleep model_PainSleep   
```

```{r, echo = FALSE}
fixed_effect <- fixef(model_PainSleep)
fixed_intercept <- fixed_effect["(Intercept)"]
fixed_slope <- fixed_effect["LastDay_PainAve_c"] ## Fixed effects

random_effects <- ranef(model_PainSleep)$ID
random_effects_df <- data.frame(
  ID = rownames(random_effects),
  RandomIntercept = random_effects[, "(Intercept)"]
) ## Random effects

random_lines <- random_effects_df %>%
  mutate(
    Intercept = fixed_intercept + RandomIntercept,
    Slope = fixed_slope
  ) ## Random intercept lines prep

sleep_range <- seq(
  from = min(df_new$LastDay_PainAve_c, na.rm = TRUE),
  to = max(df_new$LastDay_PainAve_c, na.rm = TRUE),
  length.out = 100
)  ## Define the range of predictors

prediction_df <- random_lines %>%
  crossing(LastDay_PainAve_c = sleep_range) %>%
  mutate(Today_Sleep = Intercept + Slope * LastDay_PainAve_c)

fixed_line_df <- data.frame(
  LastDay_PainAve_c = sleep_range,
  Today_Sleep = fixed_intercept + fixed_slope * sleep_range
)

ggplot() +
  geom_line(
    data = prediction_df,
    aes(x = LastDay_PainAve_c, y = Today_Sleep, group = ID),
    color = "grey",
    linetype = "dashed",
    alpha = 0.5
  ) +
  geom_line(
    data = fixed_line_df,
    aes(x = LastDay_PainAve_c, y = Today_Sleep),
    color = "black",
    size = 1
  ) +
  scale_y_continuous(
    limits = c(0, 10),
    breaks = seq(0, 10, by = 1),
    oob = scales::squish  
  ) +
  scale_x_continuous(
    limits = c(min(floor(df_new$LastDay_PainAve_c), na.rm = TRUE), 
               max(ceiling(df_new$LastDay_PainAve_c), na.rm = TRUE)),
    breaks = seq(min(floor(df_new$LastDay_PainAve_c), na.rm = TRUE), 
                 max(ceiling(df_new$LastDay_PainAve_c), na.rm = TRUE), by = 1),
    oob = scales::squish  
  ) +
  theme_minimal() +
  labs(
    title = "Random Intercept Plot: Pain vs. Sleep",
    x = "Last Day Pain (Centered)",
    y = "Today Average Sleep"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(
      size = 12,                      
      hjust = 0.45,                    
      margin = margin(b = 20)         
    ),
    axis.title.x = element_text(
      margin = margin(t = 15)         
    ),
    axis.title.y = element_text(
      margin = margin(r = 15)         
    ),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5)
  )
```

```{r}
# Obj1: Not a graph; but test/analysis of random slopes effects (for both graphs above) 
# to make sure that we have random slope effects that would justify the use of a figure showing not only bolded lines (fixed effect) but also the random slopes 
```

```{r, echo = FALSE}
model_SleepPain_rs <- lmer(Today_PainAve ~ LastDay_Sleep_c + (LastDay_Sleep_c|ID), data = df_new)

fixed_effect <- fixef(model_SleepPain_rs)
fixed_intercept <- fixed_effect["(Intercept)"]
fixed_slope <- fixed_effect["LastDay_Sleep_c"] ## Fixed effects

random_effects <- ranef(model_SleepPain_rs)$ID
random_effects_df <- data.frame(
  ID = rownames(random_effects),
  RandomIntercept = random_effects[, "(Intercept)"],
  RandomSlope = random_effects[, "LastDay_Sleep_c"]
) ## Random effects

random_lines <- random_effects_df %>%
  mutate(
    Intercept = fixed_intercept + RandomIntercept,
    Slope = RandomSlope
  ) ## Random intercept lines prep

sleep_range <- seq(
  from = min(df_new$LastDay_Sleep_c, na.rm = TRUE),
  to = max(df_new$LastDay_Sleep_c, na.rm = TRUE),
  length.out = 100
)  ## Define the range of predictors

prediction_df <- random_lines %>%
  crossing(LastDay_Sleep_c = sleep_range) %>%
  mutate(Today_PainAve = Intercept + Slope * LastDay_Sleep_c)

fixed_line_df <- data.frame(
  LastDay_Sleep_c = sleep_range,
  Today_PainAve = fixed_intercept + fixed_slope * sleep_range
)

ggplot() +
  geom_line(
    data = prediction_df,
    aes(x = LastDay_Sleep_c, y = Today_PainAve, group = ID),
    color = "grey",
    linetype = "dashed",
    alpha = 0.5
  ) +
  geom_line(
    data = fixed_line_df,
    aes(x = LastDay_Sleep_c, y = Today_PainAve),
    color = "black",
    size = 1
  ) +
  scale_y_continuous(
    limits = c(0, 10),
    breaks = seq(0, 10, by = 1),
    oob = scales::squish  
  ) +
  scale_x_continuous(
    limits = c(min(floor(df_new$LastDay_Sleep_c), na.rm = TRUE), 
               max(ceiling(df_new$LastDay_Sleep_c), na.rm = TRUE)),
    breaks = seq(min(floor(df_new$LastDay_Sleep_c), na.rm = TRUE), 
                 max(ceiling(df_new$LastDay_Sleep_c), na.rm = TRUE), by = 1),
    oob = scales::squish  
  ) +
  theme_minimal() +
  labs(
    title = "Random Slope Plot: Sleep vs. Pain",
    x = "Last Day Sleep (Centered)",
    y = "Today Average Pain"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(
      size = 12,                      
      hjust = 0.45,                    
      margin = margin(b = 20)         
    ),
    axis.title.x = element_text(
      margin = margin(t = 15)         
    ),
    axis.title.y = element_text(
      margin = margin(r = 15)         
    ),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5)
  )

######

model_PainSleep_rs <- lmer(Today_Sleep ~ LastDay_PainAve_c + (LastDay_PainAve_c|ID), data = df_new)

fixed_effect <- fixef(model_PainSleep_rs)
fixed_intercept <- fixed_effect["(Intercept)"]
fixed_slope <- fixed_effect["LastDay_PainAve_c"] ## Fixed effects

random_effects <- ranef(model_PainSleep_rs)$ID
random_effects_df <- data.frame(
  ID = rownames(random_effects),
  RandomIntercept = random_effects[, "(Intercept)"],
  RandomSlope = random_effects[, "LastDay_PainAve_c"]
) ## Random effects

random_lines <- random_effects_df %>%
  mutate(
    Intercept = fixed_intercept + RandomIntercept,
    Slope = RandomSlope
  ) ## Random intercept lines prep

sleep_range <- seq(
  from = min(df_new$LastDay_PainAve_c, na.rm = TRUE),
  to = max(df_new$LastDay_PainAve_c, na.rm = TRUE),
  length.out = 100
)  ## Define the range of predictors

prediction_df <- random_lines %>%
  crossing(LastDay_PainAve_c = sleep_range) %>%
  mutate(Today_PainAve = Intercept + Slope * LastDay_PainAve_c)

fixed_line_df <- data.frame(
  LastDay_PainAve_c = sleep_range,
  Today_PainAve = fixed_intercept + fixed_slope * sleep_range
)
```

\newpage

```{r, echo = FALSE}
ggplot() +
  geom_line(
    data = prediction_df,
    aes(x = LastDay_PainAve_c, y = Today_PainAve, group = ID),
    color = "grey",
    linetype = "dashed",
    alpha = 0.5
  ) +
  geom_line(
    data = fixed_line_df,
    aes(x = LastDay_PainAve_c, y = Today_PainAve),
    color = "black",
    size = 1
  ) +
  scale_y_continuous(
    limits = c(0, 10),
    breaks = seq(0, 10, by = 1),
    oob = scales::squish  
  ) +
  scale_x_continuous(
    limits = c(min(floor(df_new$LastDay_PainAve_c), na.rm = TRUE), 
               max(ceiling(df_new$LastDay_PainAve_c), na.rm = TRUE)),
    breaks = seq(min(floor(df_new$LastDay_PainAve_c), na.rm = TRUE), 
                 max(ceiling(df_new$LastDay_PainAve_c), na.rm = TRUE), by = 1),
    oob = scales::squish  
  ) +
  theme_minimal() +
  labs(
    title = "Random Slope Plot: Pain vs. Sleep",
    x = "Last Day Pain (Centered)",
    y = "Today Average Sleep"
  ) +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(
      size = 12,                      
      hjust = 0.45,                    
      margin = margin(b = 20)         
    ),
    axis.title.x = element_text(
      margin = margin(t = 15)         
    ),
    axis.title.y = element_text(
      margin = margin(r = 15)         
    ),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.5)
  )
```

```{r}
AIC(model_SleepPain_rs, model_SleepPain)
AIC(model_PainSleep_rs, model_PainSleep)
```

Lower AIC indicates a better model fit. Both models with random slope have lower AIC than the randomn intercept only model. This validates our use of random slope plot. 

















```{r, eval = FALSE, include = FALSE}
model_GI <- lmer(Sleep ~ LastDay_PainAve * B_Psych_PCS + (1|ID), data = df_new)
summary(model_GI)

model_GI <- lmer(AvePain ~ LastDay_Sleep * B_Psych_PCS + (1|ID), data = df_new)
summary(model_GI)

model_GI <- lmer(GlobalImprovement ~ AvePain * B_Psych_PCS + (1|ID), data = df_new)
summary(model_GI)

model_GI <- lmer(GlobalImprovement ~ B_Psych_PCS + (1|ID), data = df_new)
summary(model_GI)
```


```{r, eval = FALSE, include = FALSE}
model_sleepvsGI <- lmer(GlobalImprovement ~ Sleep * Day + (Day|ID), data = df_new)
summary(model_sleepvsGI)

model_sleepvsNextPain <- lmer(NextDay_PainAve ~ Sleep * Day + (Day|ID), data = df_new)
summary(model_sleepvsNextPain)

model_PainvsNextSleep <- lmer(NextDay_Sleep ~ AvePain * Day + (Day|ID), data = df_new)
summary(model_PainvsNextSleep)
```


```{r, eval = FALSE, include = FALSE}
## lmer Analysis (GlobImp vs. AvePain_Change)
model_1 <- lmer(GlobalImprovement ~ AvePain_Change + Day + (1|ID), data = df2_greaterthan7) # Random intercept only
summary(model_1)
model_2 <- lmer(GlobalImprovement ~ AvePain_Change + Day + (Day|ID), data = df2_greaterthan7) # Random slope + Random intercept
summary(model_2)
model_3 <- lmer(GlobalImprovement ~ AvePain_Change + Mood_Change + Day + (1|ID), data = df2_greaterthan7) # Adjust AvePain_Change and Mood_Change
summary(model_3)
model_4 <- lmer(GlobalImprovement ~ AvePain_Change + Mood_Change + Day + (Day|ID), data = df2_greaterthan7) # Add random slope
summary(model_4)

## Model comparison for above analysis
anova(model_1, model_2)
anova(model_3, model_4)

## lmer Analysis (GlobImp vs. Mood_Change)
model_5 <- lmer(GlobalImprovement ~ Mood_Change + Day + (1|ID), data = df2_greaterthan7) # random intercept only
summary(model_5)
model_6 <- lmer(GlobalImprovement ~ Mood_Change + Day + (Day|ID), data = df2_greaterthan7) # Add random slope
summary(model_6)
anova(model_5, model_6) # prefer random slope + intercept
```

```{r, eval = FALSE, include = FALSE}
model_4plus <- lmer(GlobalImprovement ~ AvePain_Change + Mood_Change + Sleep + Day + (Day|ID), 
                    data = df2_greaterthan7)
summary(model_4plus)
confint(model_4plus)
```

```{r, eval = FALSE, include = FALSE}
model_4plus_level2 <- lmer(GlobalImprovement ~ AvePain_Change + Mood_Change + Sleep + B_Demog_Age + B_Demog_Gender + B_Psych_PCS + Day + (Day|ID), 
                    data = df_new)

model_test1 <- lmer(GlobalImprovement ~ B_Demog_Age + Day + (Day|ID), 
                    data = df_new)
summary(model_test1)

model_test2 <- lmer(GlobalImprovement ~ B_Demog_Age + AvePain_Change + Mood_Change + Sleep + Day + (Day|ID), 
                    data = df_new)
summary(model_test2)

model_test3 <- lmer(GlobalImprovement ~ B_Psych_PCS + Day + (Day|ID), 
                    data = df_new)
summary(model_test3)

model_test4 <- lmer(GlobalImprovement ~ B_Psych_PCS + AvePain_Change + Mood_Change + Sleep + Day + (Day|ID), 
                    data = df_new)
summary(model_test4)

model_test5 <- lmer(GlobalImprovement ~ Mood_Change*B_Demog_Gender + Day + (Day|ID), data = df_new)
summary(model_test5)

model_test6 <- lmer(GlobalImprovement ~ B_Demog_Gender*AvePain_Change + Day + (Day|ID),
                    data = df_new)
summary(model_test6)

model_test7 <- lmer(GlobalImprovement ~ B_Demog_Age*AvePain_Change + Day + (Day|ID),
                    data = df_new)
summary(model_test7)

model_test8 <- lmer(GlobalImprovement ~ B_Demog_Age*Mood_Change + Day + (Day|ID),
                    data = df_new)
summary(model_test8)

#### PCS * AvePain_Change ####
model_test9 <- lmer(GlobalImprovement ~ B_Psych_PCS*AvePain_Change + Day + (Day|ID),
                    data = df_new)
summary(model_test9)

model_test10 <- lmer(GlobalImprovement ~ B_Psych_PCS*Mood_Change + Day + (Day|ID),
                    data = df_new)
summary(model_test10)
```



```{r, eval = FALSE, include = FALSE}
## June 20, 2024, Meeting
## Pie Chart
df_baseline_wide <- df1
gender <- factor(df_baseline_wide$B_Demog_Gender)
value <- c(sum(gender == 1), sum(gender ==2 ))
gender <- unique(gender)

pie_data <- data.frame(group = gender,
                       value = value)

ggplot(pie_data, aes(x = "", y = value, fill = gender)) +
  geom_bar(stat="identity", width = 1, color = "white") +
  coord_polar("y", start = 0) +
  theme_void()

## Bar plot shoing % of patient with specific pain conditions
df_baseline_wide_temp <- df_baseline_wide |> filter(!is.na(B_Clin_PainSite)) |> mutate(B_Clin_PainSite = factor(B_Clin_PainSite)) |> mutate(B_Clin_PainSite = case_when(
  B_Clin_PainSite == 1 ~ "Head/Face",
  B_Clin_PainSite == 2 ~ "Neck/Upper ext",
  B_Clin_PainSite == 3 ~ "Chest",
  B_Clin_PainSite == 4 ~ "Abdo/Pelvic",
  B_Clin_PainSite == 5 ~ "Low Back",
  B_Clin_PainSite == 6 ~ "Lower ext",
  B_Clin_PainSite == 7 ~ "Multiple sites"
) ) |> group_by(B_Clin_PainSite) |> summarise(count = n()) |> mutate(perc = count/sum(count))
ggplot(df_baseline_wide_temp, aes(x = B_Clin_PainSite, y = perc*100, fill = B_Clin_PainSite)) +
  geom_bar(stat = "identity") +
  scale_fill_discrete(name = "Pain Location") +
  labs(title = "Pain Location Count Data", 
       x = "Pain Location", 
       y = "% of Count") +
  theme_minimal() +
  theme(element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

## Violin Plot
data_boxplot_vars <- df_total |>
  select(c(ID, AvePain, Sleep, Mood)) |>
  group_by(ID) |>
  summarise(mean_AvePain = mean(AvePain, na.rm = TRUE),
            mean_Sleep = mean(Sleep, na.rm = TRUE),
            mean_Mood = mean(Mood, na.rm = TRUE)) |>
  pivot_longer(cols = c("mean_AvePain", "mean_Sleep", "mean_Mood"),
               names_to = "Variable",
               values_to = "Value")

data_boxplot_vars |>
  ggplot(aes(x = Variable, y = Value, fill = Variable)) +
  geom_violin() +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  scale_x_discrete(labels = c("AvePain", "Sleep", "Mood")) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) + # Remove background grids
  ggtitle("Average Score among 30 days for pain, sleep and mood") +
  xlab("Daily diary subject")
```

```{r, eval = FALSE, include = FALSE}
## Multilevel plot
library(broom.mixed)
Model_pain <- lmer(GlobalImprovement ~ AvePain + (1|ID), data = df_new)
summary(Model_pain)

new_data <- expand.grid(AvePain = seq(min(df_new$AvePain, na.rm = TRUE), max(df_new$AvePain, na.rm = TRUE), length.out = 100))
new_data$fit <- predict(Model_pain, new_data, re.form = NA)
pred <- predict(Model_pain, new_data, re.form = NA, se.fit = TRUE)
new_data$se <- pred$se.fit
new_data <- new_data %>%
  mutate(lower = fit - 1.96 * se, upper = fit + 1.96 * se)
ggplot(new_data, aes(x = AvePain, y = fit)) +
  geom_line(color = "blue") +
  #geom_jitter(data = df_new, aes(x = AvePain, y = GlobalImprovement), color = "black", size = 0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "gray", alpha = 0.5) +
  scale_y_continuous(limits=c(0, 10), breaks = seq(0, 10, 1)) +
  scale_x_continuous(limits=c(0, 10), breaks = seq(0, 10, 1)) +
  theme_minimal() +
  labs(x = "AvePain", y = "Global Improvement", 
       title = "Fitted Line and 95% CI for AvePain vs. Global Improvement") +
  theme(element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 

## Sleep
Model_sleep <- lmer(GlobalImprovement ~ Sleep + (1|ID), data = df_new)
summary(Model_sleep)

new_data <- expand.grid(Sleep = seq(min(df_new$Sleep, na.rm = TRUE), max(df_new$Sleep, na.rm = TRUE), length.out = 100))
new_data$fit <- predict(Model_sleep, new_data, re.form = NA)
pred <- predict(Model_sleep, new_data, re.form = NA, se.fit = TRUE)
new_data$se <- pred$se.fit
new_data <- new_data %>%
  mutate(lower = fit - 1.96 * se, upper = fit + 1.96 * se)
ggplot(new_data, aes(x = Sleep, y = fit)) +
  geom_line(color = "blue") +
  #geom_jitter(data = df_new, aes(x = Sleep, y = GlobalImprovement), color = "black", size = 0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "gray", alpha = 0.5) +
  scale_y_continuous(limits=c(0, 10), breaks = seq(0, 10, 1)) +
  scale_x_continuous(limits=c(0, 10), breaks = seq(0, 10, 1)) +
  theme_minimal() +
  labs(x = "Sleep", y = "Global Improvement", 
       title = "Fitted Line and 95% CI for Sleep vs. Global Improvement") +
  theme(element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

## Mood
Model_mood <- lmer(GlobalImprovement ~ Mood + (1|ID), data = df_new)
summary(Model_mood)

new_data <- expand.grid(Mood = seq(min(df_new$Mood, na.rm = TRUE), max(df_new$Mood, na.rm = TRUE), length.out = 100))
new_data$fit <- predict(Model_mood, new_data, re.form = NA)
pred <- predict(Model_mood, new_data, re.form = NA, se.fit = TRUE)
new_data$se <- pred$se.fit
new_data <- new_data %>%
  mutate(lower = fit - 1.96 * se, upper = fit + 1.96 * se)
ggplot(new_data, aes(x = Mood, y = fit)) +
  geom_line(color = "blue") +
  #geom_jitter(data = df_new, aes(x = Mood, y = GlobalImprovement), color = "black", size = 0.2) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "gray", alpha = 0.5) +
  scale_y_continuous(limits=c(0, 10), breaks = seq(0, 10, 1)) +
  scale_x_continuous(limits=c(0, 10), breaks = seq(0, 10, 1)) +
  theme_minimal() +
  labs(x = "Mood", y = "Global Improvement", 
       title = "Fitted Line and 95% CI for Mood vs. Global Improvement") +
  theme(element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

```

```{r, eval = FALSE, include = FALSE}
median_B_Psych_PCS <- median(df_new$B_Psych_PCS, na.rm = TRUE)

data_above_median <- df_new %>% filter(B_Psych_PCS > median_B_Psych_PCS)
data_below_median <- df_new %>% filter(B_Psych_PCS <= median_B_Psych_PCS)

model_above <- lmer(GlobalImprovement ~ AvePain + (1 | ID), data = data_above_median)
model_below <- lmer(GlobalImprovement ~ AvePain + (1 | ID), data = data_below_median)

new_data_above <- data_above_median %>%
  group_by(AvePain) %>%
  summarize(B_Psych_PCS = median(B_Psych_PCS), .groups = 'drop')

new_data_below <- data_below_median %>%
  group_by(AvePain) %>%
  summarize(B_Psych_PCS = median(B_Psych_PCS), .groups = 'drop')

new_data_above <- new_data_above %>%
  mutate(Predicted = predict(model_above, newdata = new_data_above, re.form = NA),
         se = sqrt(predict(model_above, newdata = new_data_above, re.form = NA, se.fit = TRUE)$se.fit))

new_data_below <- new_data_below %>%
  mutate(Predicted = predict(model_below, newdata = new_data_below, re.form = NA),
         se = sqrt(predict(model_below, newdata = new_data_below, re.form = NA, se.fit = TRUE)$se.fit))

new_data_above <- new_data_above %>%
  mutate(Lower = Predicted - 1.96 * se,
         Upper = Predicted + 1.96 * se,
         Catastrophizing = "Above Median")

new_data_below <- new_data_below %>%
  mutate(Lower = Predicted - 1.96 * se,
         Upper = Predicted + 1.96 * se,
         Catastrophizing = "Below Median")

plot_data <- bind_rows(new_data_above, new_data_below)

ggplot(plot_data, aes(x = AvePain, y = Predicted, color = Catastrophizing)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = Lower, ymax = Upper, fill = Catastrophizing), alpha = 0.2) +
  labs(title = "Association between daily pain intensity (Lev1) and \
       condition improvement (Lev1) as a function \
       of catastrophizing (Lev2)",
       x = "Daily pain intensity (Lev1; 0-10)",
       y = "Global Improvement (0-10)") +
  theme_minimal() +
  theme(element_blank()) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 

```


```{r, eval = FALSE, include = FALSE}
Model_pain_ef <- tidy(Model_pain, effects = "fixed", conf.int = TRUE)
ggplot(Model_pain_ef, aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "gray", alpha = 0.2) +
  theme_minimal() +
  labs(x = "Fixed Effects", y = "Estimate", 
       title = "Fixed Effects with 95% Confidence Intervals",
       subtitle = "Linear Mixed-Effects Model") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```










