---
title: "Martel Paper - Daily OA Pain"
date: "2024-03-18"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
urlcolor: blue
header-includes:    
  - \usepackage{lastpage}
  - \usepackage{fancyhdr}
  - \usepackage{setspace}
  - \usepackage{float}
  - \pagestyle{fancy}
  - \fancyhead[CO, CE]{Jiaqi Bi}
  - \fancyhead[LE, RO]{Martel et al. Daily OA Pain}
  - \fancyfoot[CO, CE]{\thepage \ of \pageref{LastPage}}
  - \floatplacement{figure}{H}
mainfont: Times New Roman
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling

```{r, warning = FALSE, message = FALSE}
## Load packages
library(tidyverse)
library(ggplot2)
library(tidyr)
library(haven)## This library provides functions to read sav file into R
library(lme4)
library(lmerTest)
```


```{r}
## Read data
data_paper <- read_sav("Dataset; 2024.1.sav")
checkdf3 <- data_paper |> 
  subset(ID == 2072) |> 
  select(c(ID, 
           Level1_Even_DateIn, 
           Level1_Even_TimeIn))
```

### Adjusting Wave Day ###

\begin{itemize}
  \item $DT_i$ combines $D_i$ and $T_i$: `DateTime` variable
  \item $DT_0$ is the first response `DateTime` for each patient
  \item $W_i$ is the adjusted `Wave\_Day` variable
  \item Add a grace period $G$ for calculating the adjusted $W_i$, in our case $G=6$ hours
  \item Calculate the datetime difference $H_i$ in **hours** from the first response, incorporating the grace period:
  $$
  H_i = DT_i-DT_{i-1}
  $$
  \item Then apply the grace period indicator $I_i$:
  $$
  I_i=
  \begin{cases}
  1 & \text{if } H_{i} \leq 24 + G \\
  \left\lceil\frac{H_{i} - G}{24}\right\rceil & \text{otherwise}
  \end{cases}
  $$
  \item The initial response for `Wave\_Day` is 1, i.e., $W_0=1$, then the adjusted `Wave\_Day` $W_i$ is 
  $$
  W_i=\sum_{i=0}^{i-1} I_i
  $$
\end{itemize}


```{r}
## Consecutive Days - Grace Period 6 hours
data_paper <- data_paper |>
  mutate(Lev1_DateTimeIn = as.POSIXct(strptime(paste(Level1_Even_DateIn, 
                                                     Level1_Even_TimeIn), 
                                          format="%Y-%m-%d %H:%M:"))) |>
  arrange(ID, Lev1_DateTimeIn) |>
  group_by(ID) |>
  mutate(
    TimeDiffHours = as.numeric(difftime(Lev1_DateTimeIn, 
                                        lag(Lev1_DateTimeIn, 
                                            default = first(Lev1_DateTimeIn)), 
                                        units = "hours")), # T diff
    WithinGracePeriod = if_else(TimeDiffHours <= 30, 
                                1, 
                                ceiling((TimeDiffHours - 6) / 24)), # Check grace period
    Wave_Day_Adjusted = cumsum(WithinGracePeriod) # Adjusted Wave_Day
  ) |>
  ungroup()

###### Check if the above approach is correct ######
checkdf <- data_paper |> select(c(ID, 
                                  Lev1_DateTimeIn, 
                                  TimeDiffHours, 
                                  WithinGracePeriod, 
                                  Wave_Day_Adjusted))
checkdf2 <- checkdf |> subset(ID == 2072) # Weird ID 2072
max(checkdf$Wave_Day_Adjusted, na.rm = TRUE) # Maybe these two are different patients?
```


```{r}
## Fill in the gap of Wave_Day
data_paper2 <- data_paper |>
  group_by(ID) |>
  complete(Wave_Day = 1:14) |>
  ungroup()
```

```{r}
## Check the aberrant values
summary(data_paper2$IndexLev1_NegativeAffect_Total) # Lev 1 Negative Affect 
summary(data_paper2$IndexLev1_Catastrophizing_Total) # Lev 1 Catas
summary(data_paper2$IndexLev2_QST_BaselinePPTh) # Lev 2 PPThs?
summary(data_paper2$IndexLev2_QST_TSPAve) # Lev 2 TSP?
summary(data_paper2$IndexLev2_QST_CpmTrialAve) # Lev 2 CPM?
summary(data_paper2$IndexLev1_PainAverage) # Lev 1 Pain?

###### Check if lagged value is correct ######
data_paper2_check <- data_paper2 |>
  select(c(ID, Wave_Day_Adjusted, IndexLev1_PainAverage, IndexLev1_PainAverage_Lagged))
```

$$
APE(t_i)=I(PAIN(t_i)-PAIN(t_i-1) \geq 20)
$$

```{r}
## APE index
data_paper2 <- data_paper2 |>
  group_by(ID) |>
  mutate(APE = ifelse(IndexLev1_PainAverage - IndexLev1_PainAverage_Lagged >= 20, 1, 0))

###### Check if it is correctly coded ######
data_paper2_check <- data_paper2 |>
  select(c(ID, Wave_Day_Adjusted, IndexLev1_PainAverage, IndexLev1_PainAverage_Lagged, APE))
```

For calculating the RPE based on the within person mean, define the indicator that the pain is above the average pain for person $i$ on day $t_i$.

$$
A(t_i)=I\Big(PAIN(t_i)>\frac{1}{n_t}\sum_{t=1}^{n_i}PAIN(t_i)\Big)
$$
Then define the RPE given $A(t_i)=1$ for person $i$ on day $t_i$. 

$$
RPE(t_i)=I\Big(PAIN(t_i)\leq \frac{1}{n_t}\sum_{t=1}^{n_i}PAIN(t_i)\Big)\times A(t_i-1)
$$

```{r}
## RPE Index using within person mean
data_paper2 <- data_paper2 |>
  group_by(ID) |>
  mutate(AVE = mean(IndexLev1_PainAverage, na.rm = TRUE),
         A = ifelse(IndexLev1_PainAverage > AVE, 1, 0),
         A_lag = lag(A),
         Lev1_RPE_useMean = ifelse(A_lag == 1, 
                                   ifelse(IndexLev1_PainAverage <= AVE, 1, 0), 0)) 

###### Check ######
data_paper2_check <- data_paper2 |>
  select(c(ID, Wave_Day_Adjusted, IndexLev1_PainAverage, IndexLev1_PainAverage_Lagged, 
           AVE, A, A_lag, Lev1_RPE_useMean))
# 203 RPEs
  
```

For calculating the RPE based on the APE, 

$$
RPE(t_i)=I\Big(PAIN(t_i)\leq \frac{1}{n_t}\sum_{t=1}^{n_i}PAIN(t_i)\Big)\times APE(t_i-1)
$$

```{r}
## RPE Index using APE
data_paper2 <- data_paper2 |>
  group_by(ID) |>
  mutate(APE_lag = lag(APE), 
         Lev1_RPE_useAPE = ifelse(APE_lag == 1, 
                                   ifelse(IndexLev1_PainAverage <= 20, 1, 0), 0)) 

###### Check ######
data_paper2_check <- data_paper2 |>
  select(c(ID, Wave_Day_Adjusted, IndexLev1_PainAverage, IndexLev1_PainAverage_Lagged, 
           APE, APE_lag, Lev1_RPE_useAPE))
# 4 RPEs
```






